Accounting for uncertainty in simulating greenhouse gas emissions from dairy cattle in Kenya

### 2022-06-16 Manuscript_figures.Rmd
updated milk yield measurements

The

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
all_farms_per_capita<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") + theme_bw() +
    xlab(expression(Total~greenhouse~gas~emissions~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per capita")

all_farms_per_capita

ggsave("figures/Fig_S1_emissions_from_all_farms_per_capita.png",width=9,height=5)

@@

We can now look at summary statistics of total and per capita emissions from cattle across livestock-keeping smallholder farms in Kenya. To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per capita milk yields of >500 kg of milk.

@@

emissions_summary<-rbind(data.frame("Emissions per farm"=round(quantile(GHG_simulation_scenarios$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)),
                         data.frame("Emissions per farm"=round(quantile(dairy_households$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per capita"=round(quantile(dairy_households$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)))
emissions_summary$Quantile<-c("Minimum","5% quantile", "Median", "95% quantile", "Maximum",
                              "Minimum","5% quantile", "Median", "95% quantile", "Maximum")
rownames(emissions_summary)<-NULL

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per capita" = 1, "Emissions intensity" = 1))

@@

We're interested in the suitability of total per capita milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.

Here's the relationship between these two variables, for farms with per-capita milk yields of >500 kg per year.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per capita (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b). The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig. 4b).*"}

plot1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) + 
  geom_point(shape = 20, color = "blue", size = 2) + 
  stat_smooth(method = "loess", fullrange = TRUE) +
  geom_density_2d() +
  geom_rug(color="lightskyblue3") + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})))+
  theme_bw() +
  theme(legend.position = c(0.15, 0.9))

@@

densi<-ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) +
  stat_density_2d(
    geom = "raster",
    aes(fill = after_stat(density)),
    contour = FALSE
    ) +
  scale_fill_viridis_c() + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~capita~(kg~head^{-1}~year^{-1})),
       fill="Probability \ndensity") +
  theme_bw() + geom_vline(xintercept=c(2000,3000),color=c("#F8766D","#00BFC4"),alpha=.5, lwd=2)

@@

According to these distributions, the emissions intensity at a particular per capita milk yield can be described by a slice through the probability surface.

```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-capita milk production among livestock-keeping smallholder farms in Kenya. Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)
slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=3000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)

one_slice<-ggplot(data=data.frame(CO2eq_per_milk=slice_2000$resampled), aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")

@@

two_slices<-ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~capita~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")

@@

By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other. We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per capita.

```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-capita milk production levels. Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-capita milk production and emissions intensity. This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per capita milk production and greenhouse gas emissions intensity.*"}

@@

interventions_scatter<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per capita milk yield (kg)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per capita milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}

interventions_scatter_relative<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per capita milk yield (%)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

metrics <- c("Milk yield per capita", "Emissions intensity of milk production")
names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")

@@

kable(intervention_summary, booktabs = TRUE,digits=0,
     # col.names = c("Intervention",
       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.  </i></p>") %>%
  pack_rows(index = c("Impact on milk yield per capita (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

@@

interventions_scatter_end<-ggplot(data=normal_plot,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per capita milk yield (kg)",
       color="Intervention")  +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per capita and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

interventions_scatter_relative_end<-ggplot(data=normal_plot,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per capita milk yield (%)",
       color="Intervention") +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

kable(intervention_end_summary, booktabs = TRUE,digits=0,
      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures. At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
  pack_rows(index = c("Impact on milk yield per capita (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

confidence_level_analysis$confidence<-confidence_level_analysis$confidence*100
toplot<-melt(confidence_level_analysis,id.vars = "confidence")
toplot$variable <- factor(toplot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
                             "Type I error - failure to reward 'Fewer replacement males' adopters",
                             "Type I error - failure to reward 'Calliandra feed' adopters",
                             "Type I error - failure to reward 'Balanced diet' adopters",
                             "Type II error - rewarding non-adopters"))

ggplot(toplot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")

@@

>>>>

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 4), fig.cap = "***Figure S1.*** *Greenhouse gas emissions from livestock per head of cattle across 4140 simulated livestock-keeping households in Kenya (informed by data collected on 414 farms). Plausible estimates are generated through Monte Carlo simulation, based on 1000 runs of a greenhouse gas emissions model, with uncertainty introduced by specifying inputs as probability distributions.*"}
all_farms_per_head<-ggplot(data=GHG_simulation_scenarios$y, aes(pc_on_farm)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") + theme_bw() +
    xlab(expression(Total~greenhouse~gas~emissions~per~head~(kg~CO[2]*'-eq'~kg^-1))) + ggtitle("Emissions per head")

all_farms_per_head

ggsave("figures/Fig_S1_emissions_from_all_farms_per_head.png",width=9,height=5)

@@

We can now look at summary statistics of total and per head emissions from cattle across livestock-keeping smallholder farms in Kenya. To exclude farms that do not appear to be specializing in livestock, some of which feature extremely high (up to infinity in cases where no milk is produced) emissions intensity, we also look at summary statistics across only the farms with per head milk yields of >500 kg of milk.

@@

emissions_summary<-rbind(data.frame("Emissions per farm"=round(quantile(GHG_simulation_scenarios$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per head"=round(quantile(GHG_simulation_scenarios$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(GHG_simulation_scenarios$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)),
                         data.frame("Emissions per farm"=round(quantile(dairy_households$y$on_farm,
                                                                  c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions per head"=round(quantile(dairy_households$y$pc_on_farm,
                                                                    c(0,0.05,0.5,0.95,1))/1000,1),
                                    "Emissions intensity"=round(quantile(dairy_households$y$CO2eq_per_milk,
                                                                   c(0,0.05,0.5,0.95,1)),1)))
emissions_summary$Quantile<-c("Minimum","5% quantile", "Median", "95% quantile", "Maximum",
                              "Minimum","5% quantile", "Median", "95% quantile", "Maximum")
rownames(emissions_summary)<-NULL

kable(emissions_summary[,c(4,1,2,3)], booktabs = TRUE,
      col.names = c("Distribution attribute",
                    paste0("CO","$_2$","-eq farm","$^{-1}$"),
                    paste0("CO","$_2$","-eq head","$^{-1}$"),
                    paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S1.</b> Summary statistics of simulated greenhouse gas emissions from cattle across a population of cattle-keeping smallholder farms in Kenya</i></p>") %>% pack_rows(
  index = c("Emissions per farm" = 5, "Emissions per dairy-focused farm" = 5), background="lightgray") %>%
column_spec(2:4, width = "5cm") %>% row_spec(0, align = "c") %>% 
  add_header_above(c(" " = 1, "Emissions per farm" = 1, "Emissions per head" = 1, "Emissions intensity" = 1))

@@

We're interested in the suitability of total per head milk production per year as an indicator of the greenhouse gas emissions intensity of milk production.

Here's the relationship between these two variables, for farms with per-head milk yields of >500 kg per year.

```{r, warning=FALSE, message=FALSE, fig.dim = c(8, 6), fig.cap="***Figure 3.***  *Relationship between milk yield per head (x-axis) and emissions intensity of milk production for livestock-keeping smallholder households in Kenya, expressed as a population-level scatter plot (a) and an interpolated probability density surface (b). The vertical lines in (b) illustrate slices through the density surface at productivity levels of 2000 (red) and 3000 (blue) kg milk head^-1^ year^-1^ (the corresponding distributions are shown in Fig. 4b).*"}

plot1 <- ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) + 
  geom_point(shape = 20, color = "blue", size = 2) + 
  stat_smooth(method = "loess", fullrange = TRUE) +
  geom_density_2d() +
  geom_rug(color="lightskyblue3") + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})))+
  theme_bw() +
  theme(legend.position = c(0.15, 0.9))

@@

densi<-ggplot(dairy_households$y, aes(x = pc_milk_yield, y = CO2eq_per_milk)) +
  stat_density_2d(
    geom = "raster",
    aes(fill = after_stat(density)),
    contour = FALSE
    ) +
  scale_fill_viridis_c() + 
  labs(y=expression(Emissions~intensity~of~milk~production~(kg~CO[2]*'-eq'~kg^{-1})),
       x=expression(Milk~yield~per~head~(kg~head^{-1}~year^{-1})),
       fill="Probability \ndensity") +
  theme_bw() + geom_vline(xintercept=c(2000,3000),color=c("#F8766D","#00BFC4"),alpha=.5, lwd=2)

@@

According to these distributions, the emissions intensity at a particular per head milk yield can be described by a slice through the probability surface.

```{r varkernelslicerange, message=FALSE, warning=FALSE, fig.dim = c(12, 4), fig.cap="***Figure 4.*** *Greenhouse gas emissions intensity of milk production at specific levels of per-head milk production among livestock-keeping smallholder farms in Kenya. Distributions are shown for (a) a production level of 2000 kg milk head^-1^ year^-1^, including a histogram of the random samples the distribution is based on, and (b) production levels of 2000 and 3000 kg milk head^-1^ year^-1^.*"}
slice_2000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=2000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)
slice_3000<-uncertainty::varslice_resample(in_var = dairy_households$y$pc_milk_yield,
                                 out_var = dairy_households$y$CO2eq_per_milk, 
                                 expectedin_var=3000,
                           n = 1000,
                           n_samples = 1000,
                           out_var_sampling = 1000)

one_slice<-ggplot(data=data.frame(CO2eq_per_milk=slice_2000$resampled), aes(CO2eq_per_milk)) +
geom_histogram(aes(y=..density..), colour="black", fill="white")+
 geom_density(alpha=.2, fill="#FF6666") +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~2000~kg~milk~per~head~(kg~CO[2]*'-eq'~kg^-1))) + theme_bw() + ggtitle("a) Slice through the emissions intensity distribution")

@@

two_slices<-ggplot(data=slices, aes(Values, fill=Level)) +
 geom_density(alpha=.2) +
  ylab("Relative probability") +
    xlab(expression(Emissions~intensity~at~specific~milk~yield~per~head~(kg~CO[2]*'-eq'~kg^-1))) + labs(fill='Productivity level') + theme_bw() + ggtitle("b) Two slices at different productivity levels")

@@

By resampling from these distributions, we can determine the likelihood that a value of one of these distributions is greater than one of the other. We can systematically expand this to include more comparisons, in this case at intervals of 100 kg milk yield per head.

```{r, fig.dim = c(8, 4), fig.cap="***Figure 5.*** *Probability of a particular farm B featuring lower emissions intensity of milk production than another farm A, based on observed per-head milk production levels. Estimates are derived from pairwise comparisons of slices through a population-wide generalization surface of the relationship between per-head milk production and emissions intensity. This generalization was generated through Monte Carlo simulation, based on scenarios representing observed production settings on 414 livestock-keeping smallholder farms in Kenya.*"}

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure S2.*** *Intervention impact on per head milk production and greenhouse gas emissions intensity.*"}

@@

interventions_scatter<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per head milk yield (kg)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure 6.*** *Intervention impacts on relative per head milk production and relative greenhouse gas emissions intensity, compared to a counterfactual farm with similar properties that does not implement the respective intervention.*"}

interventions_scatter_relative<-ggplot(data=interventions_normal_milk,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(aes(color=intervention),size=0.9) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per head milk yield (%)",
       color="Intervention") +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

metrics <- c("Milk yield per head", "Emissions intensity of milk production")
names(metrics) <- c("milk_change_relative","emissions_intensity_change_relative")

@@

kable(intervention_summary, booktabs = TRUE,digits=0,
     # col.names = c("Intervention",
       #             paste0("CO","$_2$","-eq farm","$^{-1}$"),
        #            paste0("CO","$_2$","-eq head","$^{-1}$"),
         #           paste0("CO","$_2$","-eq kg","$^{-1}$")),
      caption = "<p style='color:black'><i><b>Table S3.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Impacts are evaluated against a true counterfactual baseline, in which all farms are identical except for the simulated impacts of the interventions.  </i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 5, "Impact on emissions intensity (%)" = 5), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure S6.*** *Intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

@@

interventions_scatter_end<-ggplot(data=normal_plot,
                        aes(x=milk_change,y=emissions_intensity_change)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change,y=emissions_intensity_change, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y=expression(Change~'in'~emissions~intensity~(kg~CO[2]*'-eq'~kg^-1)),
       x="Change in per head milk yield (kg)",
       color="Intervention")  +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

```{r, fig.dim = c(8, 8), fig.cap="***Figure 7.*** *Relative intervention impacts on milk production per head and greenhouse gas emissions intensity, in comparison to random changes between two evaluation times. The gray dots illustrate random changes that occurred on farms that made no management changes, whereas the colored dots show changes on farms that implemented the respective interventions.*"}

interventions_scatter_relative_end<-ggplot(data=normal_plot,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative)) +
  geom_point(color="gray",size=1) +
  geom_point(data=interventions_normal_milk_end,
                        aes(x=milk_change_relative,y=emissions_intensity_change_relative, color=intervention),size=0.8) +
  scale_color_manual(values= c("breed" = intervention_colors[1],
                              "nobull" = intervention_colors[2],
                              "lessReplace" = intervention_colors[3],
                              "Calliandra" = intervention_colors[4],
                              "balance" = intervention_colors[6]),
                    labels = c("Breed upgrade", "No bulls", "Fewer males", "Calliandra feed", "Balanced diet")) +
  theme_bw() +
  ggtitle("Intervention impact on milk yield and emissions") + 
  labs(y="Relative change in emissions intensity (%)",
       x="Relative change in per head milk yield (%)",
       color="Intervention") +
  facet_wrap(~factor(intervention,
                     levels=c("breed","nobull","lessReplace","Calliandra","balance"),
                     labels=c("Breed upgrade", "Retire unproductive bulls", "Fewer replacement males", "Calliandra feed", "Balanced diet"))  ) +
  theme(legend.position = "none")

@@

kable(intervention_end_summary, booktabs = TRUE,digits=0,
      caption = "<p style='color:black'><i><b>Table S4.</b> Summary statistics of the simulated impact of farm-scale interventions aimed at reducing greenhouse gas emissions from cattle on cattle-keeping smallholder farms in Kenya. Results are based on comparison of a baseline scenario, where no farm was assumed to have implemented the interventions and an endline scenarios, where some farms adopted the mitigation measures. At both baseline and endline, variation in outcomes expected to result from random effects was simulated using Monte Carlo simulation (10 runs per farm, for 414 farms).</i></p>") %>%
  pack_rows(index = c("Impact on milk yield per head (%)" = 6, "Impact on emissions intensity (%)" = 6), background="lightgray") %>%
column_spec(2:6, width = "3cm")

@@

confidence_level_analysis$confidence<-confidence_level_analysis$confidence*100
confidence_plot<-melt(confidence_level_analysis,id.vars = "confidence")
confidence_plot$variable <- factor(confidence_plot$variable, levels = c("breed", "nobull", "lessReplace", "Calliandra", "balance", "Type2error"),
                  labels = c("Type I error - failure to reward 'Breed upgrade' adopters",
                             "Type I error - failure to reward 'Retire unproductive bulls' adopters",
                             "Type I error - failure to reward 'Fewer replacement males' adopters",
                             "Type I error - failure to reward 'Calliandra feed' adopters",
                             "Type I error - failure to reward 'Balanced diet' adopters",
                             "Type II error - rewarding non-adopters"))

ggplot(confidence_plot, aes(x=confidence,y=value)) + geom_bar(stat="identity", fill="darkolivegreen") + facet_wrap(~ variable,nrow=6) + theme_bw() +
  labs(y="Error probability (%)", x="Required confidence that emission reductions have occurred (%)")

@@

# Narrowing key uncertainties

The sensitivity analysis indicated that uncertainty about milk yields is a key driver of output variation, with unfavorable implications for using this metric as an indicator of emissions intensity. Published evidence suggests that, with some additional effort, this uncertainty can be reduced to a measurement error of approximately 15%. We can explore the impact of such measures on the suitability of milk yield per head as an indicator of emissions intensity.

```{r, eval=FALSE}
scen_upd<-scenarios

new_error<-0.15

mean_milk_4<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_4*(1-new_error)
scen_upd[which(scen_upd$Variable=="milk_yield_4"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_4*(1+new_error)

mean_milk_5<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_5*(1-new_error)
scen_upd[which(scen_upd$Variable=="milk_yield_5"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_5*(1+new_error)

mean_milk_6<-(as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="lower"),3:ncol(scen_upd)]) +
  as.numeric(scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="upper"),3:ncol(scen_upd)])) / 2
scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="lower"),3:ncol(scen_upd)]<-mean_milk_6*(1-new_error)
scen_upd[which(scen_upd$Variable=="milk_yield_6"&scen_upd$param=="upper"),3:ncol(scen_upd)]<-mean_milk_6*(1+new_error)

write.csv(scen_upd,"data/farm_scenarios_update.csv",row.names = FALSE)
